Data Exploration

### 2019-01-15 src/background_points_no_check.Rmd
minor updates - just reran some plots with newer data

```{r}
ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png, "../output/env/background_point_check/200710_ssh_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_mlp_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/200710_o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

@@

```{r}
ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/199902_ssh_back.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/199902_mlp.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/199902_temp_surface.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/199902_temp_depth.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

@@

```{r}
ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/201406_ssh.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/201406_mlp.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/201406_temp_surface.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")
dev.copy(png,"../output/env/background_point_check/201406_temp_depth.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

@@

```{r}
bck10000points_2d <- plot(x= testbk10000$longitude_, y = testbk10000$latitude_m)
dev.copy(png,"../data/env/background_point_check/bck10000points_2d.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

>>>>

```{r}
ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Sea Surface Height Above Geoid (meters)")
dev.copy(png, "../output/env/background_point_check/200710_ssh_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Mixed layer thickness (MLP) (meters)")
dev.copy(png,"../output/env/background_point_check/200710_mlp_back_no.png") # to automatically save the plot to a png AND show it inline 
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Temperature at surface (kelvin)")
dev.copy(png,"../output/env/background_point_check/200710_temp_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Temperature at sampling depth (kelvin)")
dev.copy(png,"../output/env/background_point_check/200710_temp_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = salinity_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Salinity at surface (kelvin)")
dev.copy(png,"../output/env/background_point_check/200710_salinity_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = salinity_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Salinity at sampling depth (kelvin)")
dev.copy(png,"../output/env/background_point_check/200710_salinity_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = chl_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Chlorophyll concentration at surface (mmol.m-3)")
dev.copy(png,"../output/env/background_point_check/200710_chl_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = chl_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Chlorophyll concentration at sampling depth (mmol.m-3)")
dev.copy(png,"../output/env/background_point_check/200710_chl_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = o2_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Dissolved oxygen concentration at surface (mmol.m-3)")
dev.copy(png,"../output/env/background_point_check/200710_o2_surface_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = o2_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Dissolved oxygen concentration at sampling depth (mmol.m-3)")
dev.copy(png,"../output/env/background_point_check/200710_o2_depth_back_no.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

@@

```{r}
ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Sea Surface Height Above Geoid (meters)")
dev.copy(png,"../output/env/background_point_check/199902_ssh_back.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Mixed layer thickness (MLP) (meters)")
dev.copy(png,"../output/env/background_point_check/199902_mlp.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Temperature at surface (kelvin)")
dev.copy(png,"../output/env/background_point_check/199902_temp_surface.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Temperature at sampling depth (kelvin)")
dev.copy(png,"../output/env/background_point_check/199902_temp_depth.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

@@

```{r}
ggplot(testbk10000, aes(x = ssh_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink") + labs(x = "Sea Surface Height Above Geoid (meters)")
dev.copy(png,"../output/env/background_point_check/201406_ssh.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = mlp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Mixed layer thickness (MLP) (meters)")
dev.copy(png,"../output/env/background_point_check/201406_mlp.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_surface)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")+ labs(x = "Temperature at surface (kelvin)")
dev.copy(png,"../output/env/background_point_check/201406_temp_surface.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

ggplot(testbk10000, aes(x = temp_depth)) + geom_density(na.rm = TRUE, colour = "red") + geom_density(data=testbk20000 , na.rm = TRUE, colour = "blue") + geom_density(data=testbk50000 , na.rm = TRUE, colour = "green") + geom_density(data=testbk100000 , na.rm = TRUE, colour = "orange") + geom_density(data=testbk190000 , na.rm = TRUE, colour = "pink")  + labs(x = "Temperature at sampling depth (kelvin)")
dev.copy(png,"../output/env/background_point_check/201406_temp_depth.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png

@@

```{r}
bck10000points_2d <- plot(x= testbk10000$longitude_, y = testbk10000$latitude_m, xlab = "Longitude (meters)", ylab = "Latitude (meters")
dev.copy(png,"../output/env/background_point_check/bck10000points_2d.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

### 2019-01-15 src/data_exploration.Rmd
minor updates - just reran some plots with newer data

# Overview
Exploration of all the data collected on the presence points

@@

```{r}
presence <- read.csv("../output/bio/data_aea_cell_amo_nao_bins_env_depth_cellobno_env_naocor_cels.csv", header = TRUE)
head(presence)
```

@@

```{r}
unique_depths <- unique(presence$depthlayerno)
unique_depthdordered <- sort(unique_depths) #just puts the list in oder with no NA
length(unique_depthdordered)
```
ouch - 40 layers... a job for a boxplot probably

@@

```{r}
background <- read.csv("../output/bio/background_complete_obs_cels.csv", header = TRUE)
head(background)
#test <- background[245:250, ]
#test
```

@@

```{r}
pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
pres_env_cor <- cor(pres_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(pres_env_cor, "../output/env/pres_env_corr.csv", row.names = TRUE)
pres_corrplot <- corrplot(pres_env_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
dev.copy(png,"../output/env/simple_plots/pres_envcorr.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

@@

first see which nafo zones were sampled in each year
```{r}
nafo_by_yr <- with(presence, table(year, nafo_zone))
write.csv(inst_by_yr, file = "../output/bio/nafozone_by_yr.csv")
nafo_by_yr
```

and by month
```{r}
nafo_by_mth <- with(presence, table(month, nafo_zone))
write.csv(inst_by_yr, file = "../output/bio/nafozone_by_mth.csv")
nafo_by_mth
```

>>>>

# Overview
Exploration of all the data collected on the presence points + randomly generated background points

@@

```{r}
presence <- read.csv("../output/bio/presence_points_without_envdata_relooped.csv", header = TRUE)
head(presence)
```

@@

```{r}
unique_depths <- unique(presence$depthlayerno)
unique_depthdordered <- sort(unique_depths) #just puts the list in oder with no NA
length(unique_depthdordered)
```
ouch - 39 layers... a job for a boxplot probably

@@

```{r}
background <- read.csv("../output/bio/background_complete_obs_cels.csv", header = TRUE)
head(background)
```

@@

```{r}
pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth, bottom_depth))
pres_env_cor <- cor(pres_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(pres_env_cor, "../output/env/pres_env_corr.csv", row.names = TRUE)
pres_corrplot <- corrplot(pres_env_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
dev.copy(png,"../output/env/simple_plots/pres_envcorr.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

@@

first see which nafo zones were sampled in each year
```{r}
nafo_by_yr <- with(presence, table(year, nafo_zone))
write.csv(nafo_by_yr, file = "../output/bio/nafozone_by_yr.csv")
nafo_by_yr
```

and by month
```{r}
nafo_by_mth <- with(presence, table(nafo_zone, month))
write.csv(nafo_by_mth, file = "../output/bio/nafozone_by_mth.csv")
nafo_by_mth
```
